---
output:
  word_document: default
  html_document: default
---
```{r}
library(tidyverse)
library(readr)
SingleFamilyResidenceSales <- read_csv("SingleFamilyResidenceSales.csv")
SingleFamilyResidenceRental <- read_csv("SingleFamilyResidenceRental.csv")
```

# MIS503 – Final Project
## Pitman, Connie
### Zillow Home Value Index Analysis

### Wake County Home Sales  
  
  
**a. What have been the overall trends in Wake County Home Values?** Home Values are increasing.  
**b. There were dips in home values in the past 20 years. What years did these occur?**  Dips occurred in 1997,1998, 2009, 2010, 2011 and 2012.  
**c. Based on the analysis, where would be the least expensive area to purchase home?** Zebulon   **Most expensive 
area?** Cary and Apex  
**d. Are any area home values trending down?** All areas are trending up **Is there one area that stands out compared to others?** a few areas such as Zebulon, Wendll and Knightdale are increasing slower.  

  
```{r Create a tibble}
WakeCountySales <-select(SingleFamilyResidenceSales, RegionName, State, CountyName, Metro, "1996-05","1997-05", "1998-05","1999-05","2000-05","2001-05", "2002-05", "2003-05","2004-05", "2005-05", "2006-05","2007-05", "2008-05", "2009-05","2010-05", "2011-05", "2012-05","2013-05", "2014-05", "2015-05","2016-05", "2017-05", "2018-05")
WakeCountySales<-WakeCountySales %>%
 filter(State %in% c("NC"))
WakeCountySales<-WakeCountySales %>%
 filter(CountyName %in% c("Wake County"))

```

```{r Rename Years}
WakeCountySales<-rename(WakeCountySales,"1996"="1996-05")
WakeCountySales<-rename(WakeCountySales,"1997"="1997-05")
WakeCountySales<-rename(WakeCountySales,"1998"="1998-05")
WakeCountySales<-rename(WakeCountySales,"1999"="1999-05")
WakeCountySales<-rename(WakeCountySales,"2000"="2000-05")
WakeCountySales<-rename(WakeCountySales,"2001"="2001-05")
WakeCountySales<-rename(WakeCountySales,"2002"="2002-05")
WakeCountySales<-rename(WakeCountySales,"2003"="2003-05")
WakeCountySales<-rename(WakeCountySales,"2004"="2004-05")
WakeCountySales<-rename(WakeCountySales,"2005"="2005-05")
WakeCountySales<-rename(WakeCountySales,"2006"="2006-05")
WakeCountySales<-rename(WakeCountySales,"2007"="2007-05")
WakeCountySales<-rename(WakeCountySales,"2008"="2008-05")
WakeCountySales<-rename(WakeCountySales,"2009"="2009-05")
WakeCountySales<-rename(WakeCountySales,"2010"="2010-05")
WakeCountySales<-rename(WakeCountySales,"2011"="2011-05")
WakeCountySales<-rename(WakeCountySales,"2012"="2012-05")
WakeCountySales<-rename(WakeCountySales,"2013"="2013-05")
WakeCountySales<-rename(WakeCountySales,"2014"="2014-05")
WakeCountySales<-rename(WakeCountySales,"2015"="2015-05")
WakeCountySales<-rename(WakeCountySales,"2016"="2016-05")
WakeCountySales<-rename(WakeCountySales,"2017"="2017-05")
WakeCountySales<-rename(WakeCountySales,"2018"="2018-05")

```
  
```{r}
WakeCountySales<-pivot_longer(WakeCountySales,c("1996","1997","1998", "1999", "2000","2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015","2016","2017","2018"),names_to = "YR", values_to="ZHVI")
```


```{r Scatterplot}
ggplot(WakeCountySales,mapping=aes(YR,ZHVI, color=RegionName)) +
 geom_point()+
  labs(title = "Wake County Home Sales", x= "Year", y= "Home Values (in $)") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5))+
  scale_y_continuous(name="Home Values (in $)", labels = scales::comma)



```
  
```{r bar graph sales}
ggplot(WakeCountySales,mapping=aes(YR,ZHVI,color=RegionName)) +
 geom_col()+
  labs(title = "Wake County Home Sales", x= "Year", y= "Home Values (in $)") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5))+
  scale_y_continuous(name="Home Values (in $)", labels = scales::comma)

  
```
  
### NC Rental Market  

**a. What has been the overall trend in the rental market around the state? Rental costs are increasing, Are there any cities that 
have not followed this trend?** Fayetteville rental cost remain about the same  
**b. Where is the most expensive city to rent in? Least expensive?**  Asheville  
**c. You are trying decide between Wilmington and Asheville. Which market has the lowest rent?** Wilmington  

```{r}
Rentals <-select(SingleFamilyResidenceRental, RegionName, State, "2010-11", "2011-11", "2012-11","2013-11", "2014-11", "2015-11","2016-11", "2017-11", "2018-10")
Rentals<-Rentals %>%
 filter(State %in% c("NC"))
Rentals<-Rentals %>%
 filter(RegionName %in% c("Asheville", "Charlotte","Durham","Fayetteville","Raleigh","Wilmington"))
```
  
   
```{r}
Rentals<-rename(Rentals,"2010"="2010-11")
Rentals<-rename(Rentals,"2011"="2011-11")
Rentals<-rename(Rentals,"2012"="2012-11")
Rentals<-rename(Rentals,"2013"="2013-11")
Rentals<-rename(Rentals,"2014"="2014-11")
Rentals<-rename(Rentals,"2015"="2015-11")
Rentals<-rename(Rentals,"2016"="2016-11")
Rentals<-rename(Rentals,"2017"="2017-11")
Rentals<-rename(Rentals,"2018"="2018-10")
```
  
```{r}
Rentals<-pivot_longer(Rentals,c("2010", "2011", "2012", "2013", "2014", "2015","2016","2017","2018"),names_to = "YR", values_to="ZHVI")

```
  
```{r}
ggplot(Rentals,mapping=aes(YR,ZHVI, color=RegionName)) +
 geom_point()+
  labs(title = "Rental Costs in NC Cities", x= "Year", y= "Average Rental (Cost)") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5))+
  scale_y_continuous(name="Average Rental (Cost)", labels = scales::comma)


```
  
```{r}
ggplot(Rentals,mapping=aes(YR,ZHVI,color=RegionName)) +
 geom_col()+
  labs(title = "Rental Costs in NC Cities", x= "Year", y= "Average Rental (Cost)") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5))+
  scale_y_continuous(name="Average Rental (Cost)", labels = scales::comma)

```
  
### Home Values in Select Rental Markets  
  
**a. According to the results, which market has the lowest median price (represented as horizontal bar in 
box plot)?**    The Charlotte-Concord-Gastonia market has the lowest median price  
**b. The violin plot will show density meaning the wider the plot is, the more observations occur within 
that area. Which market has the most density around the median value of homes**  The Raleigh Market has the most density around the median value of homes  

  
```{r}
NCHomeSales <-select(SingleFamilyResidenceSales, RegionName, State,  Metro, "1996-05","1997-05", "1998-05","1999-05","2000-05","2001-05", "2002-05", "2003-05","2004-05", "2005-05", "2006-05","2007-05", "2008-05", "2009-05","2010-05", "2011-05", "2012-05","2013-05", "2014-05", "2015-05","2016-05", "2017-05", "2018-05")
NCHomeSales<-NCHomeSales %>%
 filter(State %in% c("NC"))
NCHomeSales<-NCHomeSales %>%
 filter(Metro %in% c("Asheville", "Charlotte-Concord-Gastonia", "Raleigh", "Wilmington")) 

```

  
```{r}
NCHomeSales<-rename(NCHomeSales,"1996"="1996-05")
NCHomeSales<-rename(NCHomeSales,"1997"="1997-05")
NCHomeSales<-rename(NCHomeSales,"1998"="1998-05")
NCHomeSales<-rename(NCHomeSales,"1999"="1999-05")
NCHomeSales<-rename(NCHomeSales,"2000"="2000-05")
NCHomeSales<-rename(NCHomeSales,"2001"="2001-05")
NCHomeSales<-rename(NCHomeSales,"2002"="2002-05")
NCHomeSales<-rename(NCHomeSales,"2003"="2003-05")
NCHomeSales<-rename(NCHomeSales,"2004"="2004-05")
NCHomeSales<-rename(NCHomeSales,"2005"="2005-05")
NCHomeSales<-rename(NCHomeSales,"2006"="2006-05")
NCHomeSales<-rename(NCHomeSales,"2007"="2007-05")
NCHomeSales<-rename(NCHomeSales,"2008"="2008-05")
NCHomeSales<-rename(NCHomeSales,"2009"="2009-05")
NCHomeSales<-rename(NCHomeSales,"2010"="2010-05")
NCHomeSales<-rename(NCHomeSales,"2011"="2011-05")
NCHomeSales<-rename(NCHomeSales,"2012"="2012-05")
NCHomeSales<-rename(NCHomeSales,"2013"="2013-05")
NCHomeSales<-rename(NCHomeSales,"2014"="2014-05")
NCHomeSales<-rename(NCHomeSales,"2015"="2015-05")
NCHomeSales<-rename(NCHomeSales,"2016"="2016-05")
NCHomeSales<-rename(NCHomeSales,"2017"="2017-05")
NCHomeSales<-rename(NCHomeSales,"2018"="2018-05")
```
  
```{r}
NCHomeSales<-pivot_longer(NCHomeSales,c("1996","1997","1998", "1999", "2000","2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015","2016","2017","2018"),names_to = "YR", values_to="ZHVI")
NCHomeSales<-NCHomeSales %>%
 filter(Metro %in% c("Asheville","Charlotte-Concord-Gastonia",   "Raleigh", "Wilmington")) 
```
  
```{r}
NCHomeSales<-NCHomeSales %>% 
group_by(Metro)

```
  
```{r}
ggplot(NCHomeSales,mapping=aes(Metro,ZHVI)) +
 geom_violin( ) + geom_boxplot(width=0.1)+
  labs(x= "Metro", y= "ZHVI") +
  theme(axis.text.x = element_text(size=8))
```

  
### Relocation Home Value Comparison


**a. Based on your analysis, which city’s housing is most affordable?** Houston is the most affordable **Least affordable?** New York  has the least affordable housing.  
**b. Which cities saw the largest change in prices over the past 5 years?** New York and Denver had the largest change over the past 5 years **Which city has remained more consistent (i.e., no huge swings up or down in home values)?**  Houston market has been the most consistent prices.  
**c. During the market downturn in 2012, which cities were most impacted?** Chicago,New York and Denver  **Which cities have recovered?**  New York and Denver  


```{r}
NationalHomeSales <-select(SingleFamilyResidenceSales, RegionName, State,  Metro, "1996-05","1997-05", "1998-05","1999-05","2000-05","2001-05", "2002-05", "2003-05","2004-05", "2005-05", "2006-05","2007-05", "2008-05", "2009-05","2010-05", "2011-05", "2012-05","2013-05", "2014-05", "2015-05","2016-05", "2017-05", "2018-05")

NationalHomeSales<-NationalHomeSales %>%
 filter(RegionName %in% c("New York","Houston","Denver", "Chicago"))%>%
  filter(State %in% c("IL", "CO", "TX", "NY"))
```
  
```{r}
NationalHomeSales<-rename(NationalHomeSales,"1996"="1996-05")
NationalHomeSales<-rename(NationalHomeSales,"1997"="1997-05")
NationalHomeSales<-rename(NationalHomeSales,"1998"="1998-05")
NationalHomeSales<-rename(NationalHomeSales,"1999"="1999-05")
NationalHomeSales<-rename(NationalHomeSales,"2000"="2000-05")
NationalHomeSales<-rename(NationalHomeSales,"2001"="2001-05")
NationalHomeSales<-rename(NationalHomeSales,"2002"="2002-05")
NationalHomeSales<-rename(NationalHomeSales,"2003"="2003-05")
NationalHomeSales<-rename(NationalHomeSales,"2004"="2004-05")
NationalHomeSales<-rename(NationalHomeSales,"2005"="2005-05")
NationalHomeSales<-rename(NationalHomeSales,"2006"="2006-05")
NationalHomeSales<-rename(NationalHomeSales,"2007"="2007-05")
NationalHomeSales<-rename(NationalHomeSales,"2008"="2008-05")
NationalHomeSales<-rename(NationalHomeSales,"2009"="2009-05")
NationalHomeSales<-rename(NationalHomeSales,"2010"="2010-05")
NationalHomeSales<-rename(NationalHomeSales,"2011"="2011-05")
NationalHomeSales<-rename(NationalHomeSales,"2012"="2012-05")
NationalHomeSales<-rename(NationalHomeSales,"2013"="2013-05")
NationalHomeSales<-rename(NationalHomeSales,"2014"="2014-05")
NationalHomeSales<-rename(NationalHomeSales,"2015"="2015-05")
NationalHomeSales<-rename(NationalHomeSales,"2016"="2016-05")
NationalHomeSales<-rename(NationalHomeSales,"2017"="2017-05")
NationalHomeSales<-rename(NationalHomeSales,"2018"="2018-05")

```
  
```{r}
NationalHomeSales<-pivot_longer(NationalHomeSales,c("1996","1997","1998", "1999", "2000","2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012", "2013", "2014", "2015","2016","2017","2018"),names_to = "YR", values_to="ZHVI")
```
  
```{r}
NationalHomeSales<-NationalHomeSales %>% 
group_by(Metro)

```

  
```{r}
ggplot(NationalHomeSales,mapping=aes(YR,ZHVI)) +
 geom_point()+ facet_wrap(~Metro)+
  labs(title = "City Home Value Comparison", x= "Year", y= "Home Values (in $)") +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5,size=5))+
  scale_y_continuous(name="Home Values (in $)", labels = scales::comma)


```

